\documentclass{ximera}
\title{Relationship}
\begin{document}

\begin{abstract}
  Relate the Jacobian matrix and the total derivative.
\end{abstract}\maketitle

In the previous activities, we wrote some code to compute
$Df(\mathbf{p})$ and the Jacobian matrix of $f$ at the point
$\mathbf{p}$.  In this activity, we observe that these are related.

\begin{exercise}
Try running the code below.
\begin{solution}
    \begin{python}
epsilon = 0.0001
n = 3 # the dimension of the domain
m = 2 # the dimension of the codomain
def component(f,i):
  return lambda p: f(p)[i]
def partial(fi,j):
  def derivative(p):
    p_shifted = p[:]
    p_shifted[j] += epsilon
    return (fi(p_shifted) - fi(p))/epsilon
  return derivative
def jacobian(f,p):
  return [[partial(component(f,i),j)(p) for j in range(n)] for i in range(m)]
def add_vector(v,w):
  return [sum(v) for v in zip(v,w)]
def scale_vector(c,v):
  return [c*x for x in v]
def vector_length(v):
  return sum([x**2 for x in v])**0.5
def D(ff):
  def Df(p):
    f = ff
    def L(v):
      return scale_vector( 1/(epsilon),
                           add_vector( f(add_vector(p, scale_vector(epsilon, v))),
                                       scale_vector(-1, f(p)) ) )
    return L
  return Df
#
def dot_product(v,w):
  return sum([x[0] * x[1] for x in zip(v,w)])
def apply_matrix(m,v):
  return [dot_product(row,v) for row in m]
#
f = lambda p: [p[0] + p[0]*p[1], p[1] * p[2]**2]
p = [1,2,0]
v = [2,2,1]
print apply_matrix(jacobian(f,p),v)
print D(f)(p)(v)

def validator():
  # I just want them to try running this
  return True
    \end{python}
\end{solution}

In this case, we set $f(x,y,z) = (x + xy, yz^2)$, and we computed
$Df(\mathbf{p})(\vec{v})$ in two different ways.  The two different methods were close---but not exactly the same.

Why are they not exactly the same?
\begin{free-response}
  The \texttt{D} function is computing $Df(\mathbf{p})$ by comparing
  $f(\mathbf{p} + \epsilon \vec{v})$ to $f(\mathbf{p})$.

  In contrast, the \texttt{jacobian} function computes
  $Df(\mathbf{p})$ by computing partial derivatives, so we are not
  actually computing $f(\mathbf{p} + \epsilon \vec{v})$ in that case,
  but rather, $f(\mathbf{p} + \epsilon \vec{e}_i)$ for various $i$'s.

  That the way $f$ changes when wiggling each component separately has
  anything to do with what happens to $f$ when we wiggle the inputs
  together boils down to the assumption that $f$ be differentiable.
  This relationship is true infinitesimally, but here we're working
  with $\epsilon = 0.0001$, so it is not surprising that it is not
  true on the nose.
\end{free-response}

\end{exercise}

\end{document}
