\begin{document}
\section{Bilinear maps}
	
	\begin{definition}
		A \textit{bilinear map} from $B: \R^n \times \R^m \to \R^k$ is a function of two vector variables which is linear in each variable separately. 
		That is
			\begin{itemize}
				\item (Additivity in the first slot) For all $\vec{v}_1,\vec{v}_2 \in \R^n$ and $\vec{w} \in \R^m$, $B(\vec{v}_1+\vec{v}_2,\vec{w}) = B(\vec{v}_1,\vec{w})+B(\vec{v}_2,\vec{w})$
				\item (Additi\vec{v}ity in the second slot) For all $\vec{v} \in \R^n$ and $\vec{w}_1,\vec{w}_2 \in \R^m$, $B(\vec{v},\vec{w}_1+\vec{w}_2) = B(\vec{v},\vec{w}_1)+B(\vec{v},\vec{w}_2)$
				\item (Scaling in each slot) For all $c \in \R$, $\vec{v} \ in \R^n$ and $\vec{w} \in R^m$, $B(c\vec{v},\vec{w}) = B(\vec{v},c\vec{w}) = cB(\vec{v},\vec{w})$
			\end{itemize}
	\end{definition}
	
	We will mostly be focusing on bilinear maps from $\R^n \times \R^n \to \R$ (which are called \textit{bilinear forms on $\R^n$}), but we will sometimes need to work with more general bilinear maps.
	
	\begin{question}
		Confirm that the map $B:\R^n \times \R^n \to \R$ given by $B(\vec{v},\vec{w}) = \vec{v} \cdot \vec{w}$ is a bilinear form.
	\end{question}
	
	\begin{question}
		$\R^n \times \R^m$ can be identified with $\R^{n+m}$.  Is a bilinear map $\R^n \times \R^m \to \R^k$ linear when viewed as a map from $\R^{n+m} \to \R^k$?
	\end{question}
	
	\begin{question}
		Let  $B:\R^2 \times \R^3 \to R$ be a bilinear mapping, and you know the following values of $B$:
			\begin{itemize}
				\item $B(\verticalvector{1,0},\verticalvector{1,0,0}) = BLAH$
				BADBAD give action on all $6$ pairs of basis vectors.  Then ask for a formula for the map in general. 
			\end{itemize}
	\end{question}
	
	\begin{question}
		Write a bilinear map  $B: \R \times \R \to \R$.  $B(x,y) = $
	\end{question}

	\begin{question}
		Given linear maps $S : \R^n \to \R$ and $T:\R^m \to \R$ define their \textit{tensor product} to be the map 
		$S \otimes T: \R^n \times \R^m \to \R$ defined by $S \otimes T(\vec{v},\vec{w}) = S(\vec{v}) T(\vec{w})$, where 
		the RHS is the product of $S(\vec{v})$ and $T(\vec{w})$  as real numbers.
		
		Prove that $S \otimes T$ is bilinear.
	\end{question}
	
	Let us pause for a minute to develop some notation.
	
	Let $e_i$ be the standard basis for $\R^n$.  Then any covector can be written as a linear combination of the covectors $e_i^\top$ (This is just saying any row 
	vector $\begin{bmatrix} a_1 & a_2  &... &a_n\end{bmatrix} = a_1\begin{bmatrix} 1& 0&0&...&0\end{bmatrix} + a_2\begin{bmatrix} 0& 1&0&...&0\end{bmatrix} 
	+ ... + a_n \begin{bmatrix} 0& 0&0&...&1\end{bmatrix}$).   The coordinate functions $\pi_i:\R^n \to \R$ given by $\pi_i (x_1,x_2,x_3,...,x_n) = x_i$ have 
	derivatives $D\pi_i = e_i^\top$.  For this reason we will write $dx_i$ as another notation for the covector $e_i^\top$.
	
	\begin{question}
		What is $dx_2(\verticalvector{3,6,4})$?
	\end{question}
	
	\begin{question}
		What is $dx_1 \otimes dx_2 (\verticalvector{2,5,3},\verticalvector{7,9,4})$?  What about $dx_2 \otimes dx_1(\verticalvector{2,5,3},\verticalvector{7,9,4})$?
	\end{question}
	
	\begin{question}
		Prove that any bilinear map $\R^n \times \R^k \to \R$ can be written uniquely in the form 
		\[
			\sum_{i=1,j=1}^{i=n,j=k} a_{ij} dx_i \otimes dx_j
		\]
		
		for some scalars $a_{ij} \in \R$
	\end{question}
	
	\begin{question}
		Write the dot product on $\R^2$ in the form given by the question above.
	\end{question}
	
	\begin{example}
		The bilinear form $\eta = -dx_1 \otimes dx_1 +dx_2 \otimes dx_2 + dx_3\otimes dx^3 + dx_4 \otimes dx_4 $ on $\R^4$ is incredibly important to physics.  
		It is called the \href{http://en.wikipedia.org/wiki/Minkowski_space}{Minkowski inner product}, and is one of the basic structures underlying the local geometry of our 
		universe.
	\end{example}
	


		
	
	
	
	
	
\end{document}