\documentclass{ximera}
\title{Multiplying matrices using dot products}
\begin{document}

\begin{abstract}
	There is a quick way to multiply matrices using dot products
\end{abstract}\maketitle

\begin{question}
	Let $M =  \begin{bmatrix}2&3\\4&5\\1&2\end{bmatrix}$, and $\vec{e}_2 = \verticalvector{0\\1\\0}$.
	\begin{solution}
	\begin{hint}
		\begin{align*}
			\vec{e}_2^\top M &= \begin{bmatrix} 0\\1\\0\end{bmatrix} \begin{bmatrix}2&3\\4&5\\1&2\end{bmatrix}\\
			&= \begin{bmatrix} 4&5\end{bmatrix}
		\end{align*}
	\end{hint}
		$\vec{e}_2^\top M$=
		\begin{matrix-answer}[name=w]
			correctMatrix = [['4','5']]
		\end{matrix-answer}
	\end{solution}
	
	Did you notice how multiplying by $\vec{e}_2^\top$ on the right selected the $2^{nd}$ row of $M$?
\end{question}

Prove that if $M$ is an $m \times n$ matrix and $e_j \in \R^m$ is the $j^{th}$ standard basis vector of $\R^m$, then
$\vec{e_j}^\top M $ is the $j^{th}$ row of $M$.
\begin{free-response}
We know that $\vec{w} = \vec{e_j}^\top M$ is a covector (row) just by looking at dimensions.  What is the $i^{th}$ entry of this row?
Well, we can only figure that out by applying the map to the basis vectors.  $\vec{e_j}^\top M\vec{e_i}$ is the dot product of $\vec{e_j}$ with the
$i^{th}$ column of $M$.  But that just selects the $j^{th}$ element of that column.  So the $i^{th}$ element of $\vec{w}$ is the $j^th$ element of the $i^{th}$ column
of $M$.  This just says that $\vec{w}$ is the $j^th$ column of $M$. (Whew.)
\end{free-response}

Now we can use this observation to great effect.  If $M$ is an $m \times n$ matrix,  $\vec{e}_j$ is the standard basis of $\R^m$ and $\vec{b}_k$ is the
standard basis of $\R^n$, then we can select $M_{j,k}$ by performing the operation $\vec{e}_j^\top M \vec{b}_k$.  
This is so important we will label it as a theorem:

\begin{theorem}
	If $M$ is an $m \times n$ matrix,  $\vec{e}_j$ is the standard basis of $\R^m$ and $\vec{b}_k$ is the
standard basis of $\R^n$, then $M_{j,k}=\vec{e}_j^\top M \vec{b}_k$.  
\end{theorem}

\begin{proof}
	The proof is simply that $M\vec{b}_k$ is by definition the $k^{th}$ column of the matrix, and by our observation above $\vec{e}_j^\top M \vec{b}_k$
	must be the $j^{th}$ row of that column vector, which consists of the single number $M_{i,j}$
\end{proof}

\begin{question}
	Let $M = \begin{bmatrix} 4&1&-2\\3&1&0\end{bmatrix}$.  
	\begin{solution}
		\begin{hint}
			By the above theorem, it will be the entry in the $2^{nd}$ row and the $1^{st}$ column of $M$
		\end{hint}
		\begin{hint}
			$\begin{bmatrix} 0&1\end{bmatrix} M \verticalvector{1\\0\\0}=3$
		\end{hint}
	$\begin{bmatrix} 0&1\end{bmatrix} M \verticalvector{1\\0\\0}=$\answer{3}
	\end{solution}
\end{question}

The philosophical import of this theorem is that we can probe the inner structure of any matrix with simple row and column vectors to find out every
component of the matrix.  What happens when we apply this insight to a product of matrices?

\begin{question}
	Let $A = \begin{bmatrix} -1 &1\\2&2\\3&0 \end{bmatrix}$ and $B = \begin{bmatrix}\end{bmatrix}$.  Let $C = AB$.
	\begin{solution}
		\begin{hint}
			By the theorem above, $C_{2,3} = \begin{bmatrix} 0&1&0\end{bmatrix} C \verticalvector{0\\0\\1\\0}$
		\end{hint}
		\begin{hint}
			So $C_{2,3} = \begin{bmatrix} 0&1&0\end{bmatrix} AB \verticalvector{0\\0\\1\\0}$
		\end{hint}
		\begin{hint}
			But $\begin{bmatrix} 0&1&0\end{bmatrix} A$ is the $2^{nd}$ row of $A$, and $B \verticalvector{0\\0\\1\\0}$ is the $3^{rd}$ column of
			$B$
		\end{hint}
		\begin{hint}
			So $\begin{bmatrix} 0&1&0\end{bmatrix} A = \begin{bmatrix} 2&2\end{bmatrix}$ and $B \verticalvector{0\\0\\1\\0} = \verticalvector{1\\9}$
		\end{hint}
		\begin{hint}
			Thus $C_{2,3} = \begin{bmatrix} 2&2\end{bmatrix} \verticalvector{1\\9} = 2(1)+2(9)=20$ 
		\end{hint}
		Without computing the whole matrix $C$, can you find
	
		$C_{2,3} = $ \answer{20}
		
	\end{solution}
	
	Wow!  So it looks like we can find the entries of a product of two matrices just by looking at the dot product of rows of the first matrix with columns of
	the second matrix!
\end{question}

\begin{theorem}
	Let $A$ and $B$ be composable matrices.  Let $C=AB$.  Then $C_{i,j}$ is the product of the $i^{th}$ row of $A$ with the $j^{th}$ column of $B$
\end{theorem}

Prove this theorem
\begin{free-response}
	We can prove this by combining the other two theorems in this section.  $C_{i,j} = \vec{e_i}^\top C \vec{e}_j$ by the second theorem.  But $C = AB$, so we have 
	$C_{i,j} = \vec{e_i}^\top AB\vec{e}_j$.  By the first theorem $\vec{e_i}^\top A$ is the $i^{th}$ row of $A$, and by our definition of matrix multiplication, 
	$B\vec{e}_j$ is the $j^{th}$ column of $B$. So $C_{i,j}$ is the product of the $i^{th}$ row of $A$ with the $k^{th}$ column of $B$.
\end{free-response}

Now try multiplying some matrices of your choosing using this method.  This is likely the definition of matrix multiplication you learned in high school (or the same thing
defined by some messy formula with a $\sum$).  Do you prefer this method?  Or do you prefer whatever method you came up with on your own earlier?  
Maybe they are the same!

Another note:  it is interesting that we are feeding two vectors $e_i$ and $e_j$ into the matrix and getting out a number somehow.  
In week $4$ we will learn that we are treading in deep water here:  this is the very tip of the iceberg of bilinear forms, which are a kind of $2$-tensor.
  
	
\end{document}
