\begin{document}
\section{Linear maps, revisited}

\begin{definition}
	Let $V$ and $W$ be two vector spaces.  A function $L: V \to W$ is called \textit{linear} if 
	\begin{itemize}
		\item (Respects vector addition) For all $\vec{v_1},\vec{v_2} \in V$, $L(v_1+v_2) = L(v_1)+L(v_2)$
		\item (Respects scalar multiplication) For all $c \in \R$ and $\vec{v} \in V$, $L(c\vec{c}) = cL(\vec{v})$
	\end{itemize}
\end{definition}

\begin{question}
	Let $V$ be the space of all polynomials in one variable $x$, and $W = \R$.  For each real number $c \in \R$, define the function
	$Eval_c: V \to W$ defined by $Eval_c(p) = p(c)$.  Show that $Eval_c$ is a linear map. 
\end{question}

\begin{question}
	Let $V$ be the set of all solutions to the differential equation $y''=-y$.  Show that $V$ is a vector space with function addition and multiplication.
	Show that the map $L:V \to V$ given by $L(f)(x)  = f(-x)$ is a linear map. 
	(In particular, make sure it really does take solutions of the differential equation to other solutions!)
	This is the "time invariance of the wave equation".
\end{question}

\begin{question}
	Let $V$ be the set of all differentiable functions on $\R$.  Consider the map $L:V \to \R$ defined by $L(f) = f'(2)$.  Show that $L$ is linear.
\end{question}

To make sure linear maps work they way we expect them to in this new context, and to flex our brains a little bit, let's prove some facts about linear functions:

\begin{question}
	Let $L:V \to W$ be a linear map.  Show that $L(\vec{0}) = \vec{0}$. 
	(Note: $\vec{0}$ means different things on either side of the equation.  On the LHS it means the additive identity of $V$, while on the RHS it means the 
	additive identity of $W$).
\end{question}

\begin{question}
	Let $V$ and $W$ be vector spaces, and define a function $\mathrm{Zero}: V \to W$ by $\mathrm{Zero}(\vec{v}) = \vec{0}$ for all $\vec{v} \in V$.  Show that 
	$\mathrm{Zero}$ is a linear function. 
\end{question}

\begin{question}
	Let $L:V \to W$ be a linear map. Show that $L(-v) = -L(v)$ for each $v\in V$
\end{question}

\begin{question}
	Let $L:V \to W$ be a linear map.  Let $N(L) = \{ \vec{v} \in V : L(\vec{v}) = \vec{0}\}$.  Show that $N(L)$ is a vector space with addition and scalar multiplication
	inherited from $V$.  It is called the ``Null space'' of $L$.
\end{question}

\begin{question}
	Let $L:V \to W$ be a linear map.  Let $Im(L) = \{ \vec{w} \in W : \text{there is a }\vec{v} \in V \text{with } L(\vec{v}) = \vec{w}\}$. 
	 Show that $Im(L)$ is a vector space with addition and scalar multiplication
	inherited from $W$.  It is called the ``Image' of $L$.
\end{question}



\end{document}