\documentclass{ximera}

\title{Span}

\begin{document}

\begin{abstract}
  Vectors can be combined; all those combinations form the ``span.''
\end{abstract}

\begin{definition}
  We say that a vector $w$ is a \textit{linear combination} of the vectors $v_1,v_2,v_3,...,v_k$ if there are scalars $a_1,a_2,...a_k$ 
  so that $w = a_1v_1+a_2v_2+...+a_kv_k$
\end{definition}
  		
\begin{definition}
  The \textit{span} of a set of vectors $v_1,v_2, ..,v_k \in \R^n$ is the set of all linear combinations of the vectors.
  Symbolically, $\text{span}(v_1,v_2,...,v_k) = \{ a_1v_1+a_2v_2+...+a_kv_k : a_1,a_2, ...,a_k \in \R\}$
\end{definition}

\begin{example}
  The span of $\verticalvector{1\\0\\0}$, $\verticalvector{0\\1\\0}$ is all vectors of the form $\verticalvector{x\\y\\0}$ for some $x,y \in \R$
\end{example}

\begin{example}
  $\verticalvector{8\\13}$ is in the span of $\verticalvector{2\\3}$ and $\verticalvector{4\\7}$ because
  $2\verticalvector{2\\3} + \verticalvector{4\\7} = \verticalvector{8\\13}$ 
\end{example}

\begin{question}
  Is $\verticalvector{3\\4\\2}$ in the span of $\verticalvector{1\\2\\0}$ and $\verticalvector{3\\-3\\0}$?
  \begin{solution}
    \begin{hint}
      The linear combinations of $\verticalvector{1\\2\\0}$ and $\verticalvector{3\\-3\\0}$ are all the vectors of the form
      $a\verticalvector{1\\2\\0} + b\verticalvector{3\\-3\\0}$ for scalars $a,b\in \R$.  Could $\verticalvector{3\\4\\2}$ be written in such a form?
    \end{hint}
    \begin{hint}
      No, because the last coordinate of all of these vectors is $0$.  In fact, graphically, the span of these two vectors is just the entire $xy$-plane, 
      and $\verticalvector{3\\4\\2}$ lives off of that plane.
    \end{hint}
    \begin{multiple-choice}
      \choice{Yes, it is in the span of those two vectors.}
      \choice[correct]{No, it is not in the span of those two vectors.}
    \end{multiple-choice}
  \end{solution}
\end{question}

Graphically, we should think of the span of one vector as the line
which contains the vector (unless the vector is the zero vector, in
which case its span is just the zero vector).
  
The span of two vectors which are not in the same line is the plane
containing the two vectors.

The span of three vectors which are not in the same plane is the
``$3$D-space'' which contains those $3$ vectors.
  	 	 
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
